---
title: "Intro to Maps"
author: "Nicole Ayala"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

# LEARNING HOW TO CREATE A MAP

Today we are going to leanr how to create a map


# Load Libraries
```{r}
library(tidyverse)
library(here)
library(maps)
library(mapdata)
library(mapproj)
```

# Inspect the data
```{r}

popdata<-read.csv(here("Week_07","Data","CApopdata.csv")) # Read in data on population in California by county

stars<-read.csv(here("Week_07","Data","stars.csv")) # read in data on number of sea stars at different field sites

world<-map_data("world") # get data for the entire world
head(world)

usa<-map_data("usa") # get data for the USA
head(usa)

italy<-map_data("italy") # get data for italy
head(italy)

states<-map_data("state") 
# get data for states
head(states)

counties<-map_data("county")

```

## Structure of the Data 

### 1. Longitude
 - Longitude is long.
 - Things to the west of the prime meridian are negative

### 2. Latitude
  - Latitude is lat.
  - Measures left to right
  
### 3. Order
  - This shows which order ggplot should connect the dots

### 4. Region and Subregion
  - Tells what region or subregion a set of points surrounds

### 5. Group
  - This is very important! 
  - ggplot2’s functions can take a group argument which controls (amongst other things) whether adjacent points should be connected by lines. If they are in the same group, then they get connected, but if they are in different groups then they don’t. Essentially, having to points in different groups means that ggplot “lifts the pen” when going between them.

# Make a Map of the World
```{r}
ggplot()+
  geom_polygon(data = world, 
               aes(x = long, # x is always LONGITUDE
                   y = lat, # y is always LATITUTDE
                   group = group, # ALWAYS SET GROUP = GROUP
                   fill = region),
               color = "plum3") +
  guides(fill = FALSE) + # to not get hundreds of colors 
  theme_light() + # makes it look simple and clean with white grid lines
  theme(panel.background = element_rect(fill = "darkblue")) +
   coord_map(projection = "mercator",
   #coord_map(projection = "sinusoidal",
            xlim = c(-180,180))
         

  
  
```




# If You Forgot About group = group
```{r}
ggplot()+
  geom_polygon(data = world, aes(x = long, y = lat))
```
# How to Make a Map of Just California
```{r}

# Time to look at the county data
head(counties)[1:3,] # only showing the first 3 rows for space                  
head(popdata) # Look at the county data

# Need to add multiple layers of data

CApopcounty<-popdata %>%
  select("subregion" = County, Population)  %>% # rename the county col
  inner_join(counties) %>%
  filter(region == "california") # some counties have same names in other states
head(CApopcounty)

# Add a Layer of Points using geom_point
# when building up with datasets please start with the map before adding any other points 

head(states) # Use the states dataset 


# Make a map :)
CA_data<-states %>%
  filter(region == "california") 

ggplot()+
geom_polygon(data = CApopcounty, # use the specific data set
             aes(x = long, # longitude
                 y = lat,  # latitude
                 group = group, # DONT FORGET THIS
                 fill = Population),
             color = "orange") +
   geom_point(data = stars, # add a point at all my sites
             aes(x = long,
                 y = lat,
                 size = star_no))+ # make size noticeable, color is also possible
  coord_map()+
  theme_light() +
  scale_fill_gradient(trans = "log10") +
  labs(size = "# stars/m2")
 
  ggsave(here("Week_07","Output","CAMap.pdf"))


 

```



  