---
title: "Class Notes Functional Programming"
author: "Nicole Ayala"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```
 
### Functional Programming 
Purpose: when to write a function, how to make a custom function
When to Use: anytime you have a task that you want to use more than once, keeps your script clean, avoids mistakes with copy and paste, minimizes places where you need to update your code


### Example
(value - min/(max - min))

```{r}
### load libraries
library(tidyverse)

df <- tibble::tibble(
  mean =5, n=10,
  StDev =2,
  a = rnorm(10), # draws 10 random values from a randomly normal distribution
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
head(df)
# mean of 0, StDev at 1 which can be altered i.e.
```
### Rescale Every Column Individually
```{r}

######
rescale01 <- function(x) {
  value<-(x-min(x, na.rm = TRUE))/(max(x, na.rm = TRUE)-min(x, na.rm = TRUE))
  return(value)
}
df %>%
  mutate(a = rescale01(a),
         b = rescale01(b),
         c = rescale01(c),
         d = rescale01(d))
```
### 3 Steps to Creating a New Function
- You need to pick a name for the function. Here I’ve used *rescale01* because this function rescales a vector to lie between 0 and 1.

- You list the inputs, or *arguments*, to the function inside function. Here we have just one argument. If we had more the call would look like function(x, y, z).

- You place the code you have developed in *body of the function*, a { block that immediately follows function(...).

- *Return* tells us what values we want returned when you run the function

```{r}
rescale01 <- function(x) {
  value<-(x-min(x, na.rm = TRUE))/(max(x, na.rm = TRUE)-min(x, na.rm = TRUE))
  return(value)
}
```
### Making a fucntion to convert F to C
Formula: temp_C <- (temp_F - 32) * 5 / 9
```{r}

# - Step 1: Name the function
F_to_C <- function() {
}
# - Step 2: Put in the equation
F_to_C <- function() { 
  temp_C <- (temp_F - 32) * 5 / 9
}
# Step 3: Decide what the arguments are

F_to_C <- function(temp_F) {
  temp_C <- (temp_F - 32) * 5 / 9 
}
# Step 4: Decide what is being returned

F_to_C <- function(temp_F) { 
  temp_C <- (temp_F - 32) * 5 / 9 
  return(temp_C)
}
# Step 5: Test it

F_to_C(32)
## [1] 0
F_to_C(212)
## [1] 100

```
```{r}
### Making a function to convert C to K
# Formula: celcius + 273.15

# - Step 1: Name the function
C_to_K <- function() {
}

# - Step 2: Put in the equation
C_to_K <- function() { 
  temp_K <- (temp_C + 273.15)
}
# -Step 3: Decide what the arguments are

C_to_K <- function(temp_C) { 
  temp_K <- (temp_C + 273.15)
}
# - Step 4: Decide what is being returned

C_to_K <- function(temp_C) { 
  temp_K <- (temp_C + 273.15)
  return(temp_K)
}

C_to_K(32)
C_to_K(273.15)
C_to_K(0)
```
### Making Plots Into a Function
```{r}
# load libraries 
library(palmerpenguins)
library(PNWColors) # for the PNW color palette 
library(rlang)
library(ggplot2)

myplot<-function(data = penguins, x = body_mass_g, y = bill_length_mm){ 
pal<-pnw_palette("Lake",3, type = "discrete") # my color palette 
ggplot(data = penguins, aes(x = {{body_mass_g}}, y = {{bill_length_mm}} , color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
}

myplot<-function(data = penguins, x, y){
pal<-pnw_palette("Lake",3, type = "discrete") # my color palette 
ggplot(data, aes(x = {{x}}, y = {{y}} , color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
}

myplot(x = body_mass_g, y = flipper_length_mm)+
  labs(x = "Body mass (g)",
       y = "Flipper length (mm)")

```
### If Else Statements
- Suppose that if a > b then f should be = to 20, else f should be equal to 10. Using if/else we:
```{r}
a <- 4
b <- 5

if (a > b) { # my question
  f <- 20 # if it is true give me answer 1
    } else { # else give me answer 2
  f <- 10
  f
}
```
### New Arguments in Plot

```{r}
myplot<-function(data = penguins, x, y ,lines=TRUE ){ # add new argument for lines
pal<-pnw_palette("Lake",3, type = "discrete") # my color palette 
ggplot(data = penguins, aes(x = {{ body_mass_g}}, y = {{bill_length_mm}} , color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
}
```

```{r}
myplot<-function(data = penguins, x, y, lines=TRUE ){ # add new argument for lines
pal<-pnw_palette("Lake",3, type = "discrete") # my color palette 
if(lines==TRUE){
ggplot(data, aes(x = {{body_mass_g}}, y = {{bill_length_mm}} , color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
}
else{
ggplot(data, aes(x = {{x}}, y = {{y}} , color = island))+
  geom_point()+
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
}
}

myplot(x = body_mass_g, y = flipper_length_mm, lines = FALSE)
myplot(x = body_mass_g, y = bill_length_mm, lines = FALSE)

```

