
title: "HW Lab Assignment"
author: "Nicole Ayala"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "LabAssignment8b"
author: "Nicole Ayala"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

### Lab Assignment for Week 8b: Functional programming

- Step 1: Name the function
- Step 2: Put in the equation
- Step 3: Decide what the arguments are
- Step 4: Decide what is being returned

```{r}
### Load in the Library

library(tidyverse)
library(here)
library(hrbrthemes) 
library(rlang)
library(ggplot2)
```

# Here is my First Function
```{r}
### Load in the Data
auto <- read.csv(here("Week_09", "Data", "auto-mpg.csv")) # one of my datasets

cars <- mtcars # my dataset
#View(mtcars)

# I will be converting miles per gallon to kilometers per liter as this was a great way to measure that a vehicle can travel while consuming a certain amount of liters of gas per km (one liter of gas) km/L is a measurement of fuel economy
mpgtokmpl <- function(mpg) {
kmpl <- (mpg * 0.425144) #  this function converts a value in miles per gallon to kilometers per liter
  return(kmpl) # return the kmpl value
}
```

# Here is my Second Function
```{r}
kmplcars <- cars %>% # piping mtcars to new name showing that I will be converting m/g to km/l with function created
  mutate(kmpl = mpgtokmpl(mpg)) # mutating name to mpgtokmpl from kmpl

kmplauto <- auto %>%
  mutate(kmpl = mpgtokmpl(mpg))

View(mtcars)

```

### Making a Plot Function


```{r}
ggplot(data=auto,aes(x=mpg,# miles per gallon, y = acceleration
           )) + #  fill by color
  geom_density(adjust=1.5, alpha=0.5,  show.legend = FALSE) + # adjusting size
  theme_ipsum_pub() + # cute rainbow theme
  theme_minimal() +
  labs(x = "Miles per Gallon", y = "Acceleration", title = "Correlation of the Acceleration within Cars and their Amount of Gallon Used", subtitle = " None") + # title and subtitles + x and y axes
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  scale_color_brewer(palette = "Blues") +
  theme(axis.title = element_text(size = 30, color = "mediumorchid4"),
  panel.background = element_rect(fill = "white"))
  # background color is white
```

