---
title: "Bad Plot vs. Good Plot"
author: "Nicole Ayala"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### **Purpose of this Assignment**
- To assess my coding ability ad knowledge of what makes a good plot versus a bad plot.

#### ~ Requirements ~
- All plots must be made in ggplot2, with code visible in the Rmarkdown document.
- The data must be visible.
- The good plot and the bad plot should be based on the same, or similar, data.
- Will be posted on Twitter
- Clearly articulate each principle of data visualization that is broken or not

## BAD PLOT




```{r cars}
### Load the libraries ###
library(tidyverse)
library(here)
library(ggplot2)
library(waffle)
library(devtools)
library(dplyr)

### Load in the Data ###
cancer<-read_csv(here("Week_08","Data", "breastcancer_UMC.csv"))
glimpse(cancer)
# View(cancer)

### Filter the Data ###
### haha jk




```
```{r}

ggplot(cancer,
       aes(x=`deg-malig`,
           fill= class),
       color = "green") +
  geom_bar(position='dodge') + 
  theme_void() +
  labs(title='Plot With Collected Data of Deg',
       x='Degree of Malignancy',
       y='Count') +
  scale_fill_manual(values = beyonce_palette(1)) +
  theme(axis.title = element_text(size = 14, color = "orange",
        panel.background = element_rect(fill = "white")
                
ggsave(here("Week_08","Output","BadPlot.png"))

```


## GOOD PLOT

```{r}

### Load the libraries ###

library(tidyverse)
library(here)
library(ggplot2)
library(ggridges)
library(ggthemes)
library(hrbrthemes)
library(beyonce)
library(janitor)
library(devtools)
library(dplyr)
library(tidyr)
library(waffle)
library(qwraps2)
library(gridExtra)
library(arm)
library(car)
library(psych)
library(caTools)
library(caret)
library(keras)


### Load in the Data ###
cancer <- read_csv(here("Week_08","Data", "breastcancer_UMC.csv"))
glimpse(cancer) # look at data
# View(cancer)
attach(cancer) # all the data variables
head(cancer) # shows beginning of  dataset
str(cancer) # shows structure of dataset
#summary(cancer) # shows each column and other attributes


# pair up corr. coefficient btwn variables
pairs.panels(cancer) # menopause and age is stat sig 
df <- data.frame(`tumor-size`=c('0-4', '5-9', '10-14', '15-19', '20-24', '25-29', '30-34', '35-39', '40-44', '45-49', '50-54'))


### Filtering the Data ### Part 1
# cancer = filter(cancer, `breast-quad`!= '?' & node.caps != '?')

### Filter Data ### Part 2
data <- cancer %>%
filter(`tumor-size`>="30-34") %>%
arrange()

#select('tumor_size', age, class, `deg-malig`) %>%
#group_by(class) %>%
#mutate(avg = mean(`tumor-size`, na.rm = TRUE)) %>%
#filter(class == "recurrence-events") %>%
#filter('tumor-size'>"10-14" & age<"60-69")

```
```{r}
# Make the Ridgeline Plot
data <- cancer %>%
filter(`tumor-size`>="30-34")
ggplot(data=data,
       aes(x= `deg-malig`, # degree of malignancy
           group = `tumor-size`, # group by tumor size
          fill=`tumor-size`, # focus on tumor sizes
          `tumor-size`= "Tumor Size")) + # wanted to try to change name 
  geom_density(adjust=1.5, alpha=0.4) + # adjusting size
  scale_fill_discrete(breaks=c('30-34', '35-39', '40-44', '45-49', '50-54')) + # arranged order of legend
  theme_ipsum_pub() + # cute rainbow theme
  labs(x = "Degree of Malignancy", y = "Density", title = "Correlation of the Degree of Malignancy w/ Tumor Size", subtitle = "There are 5 ranges of tumor sizes from 30-54 (cm)") + # title and subtitles + x and y axes
  scale_fill_manual(values = beyonce_palette(8)) +
  theme(axis.title = element_text(size = 14, color = "darkseagreen"),
        panel.background = element_rect(fill = "white"))# color the background white

ggsave(here("Week_08","Output","GoodPlot.png"), # saved the good plot
        width = 10, height = 8) # adjust size of graph in inches

```

