---
title: "Bad Plot vs. Good Plot"
author: "Nicole Ayala"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### **Purpose of this Assignment**
- To assess my coding ability ad knowledge of what makes a good plot versus a bad plot.

#### ~ Requirements ~
- All plots must be made in ggplot2, with code visible in the Rmarkdown document.
- The data must be visible.
- The good plot and the bad plot should be based on the same, or similar, data.
- Will be posted on Twitter
- Clearly articulate each principle of data visualization that is broken or not

## BAD PLOT




```{r cars}
### Load the libraries ###
library(tidyverse)
library(here)
library(ggplot2)
library(waffle)
library(devtools)
library(dplyr)

### Load in the Data ###
cancer<-read_csv(here("Week_08","Data", "breastcancer_UMC.csv"))
glimpse(cancer)
# View(cancer)

### Filter the Data ###
### haha jk


```
```{r}

ggsave(here("Week_08","Output","BadPlot.png"))

```


## GOOD PLOT

```{r}

### Load the libraries ###

library(tidyverse)
library(here)
library(ggplot2)
library(ggridges)
library(ggthemes)
library(janitor)
library(devtools)
library(lubridate)
library(dplyr)
library(colorspace)


### Load in the Data ###
cancer<-read_csv(here("Week_08","Data", "breastcancer_UMC.csv"))
glimpse(cancer)
# View(cancer)

### Filtering the Data ###
filtercancer<-cancer %>%
clean_names %>%
filter(class != 'false recurrence events') %>%
select('tumor_size', age, class, breast ) %>%
group_by('tumor-size', breast) %>%
mutate(avg = mean(age, na.rm = TRUE)) %>%
filter(class == "recurrence-events") %>%
filter('tumor-size'>"10-14" & age<"60-69")

```
```{r}
# Make the Ridgeline Plot
filtercancer %>%
  ggplot( aes(x=tumor_size)) +
    geom_density(fill="#69b3a2", color="#e9ecef", alpha=0.8) +
  scale_fill_brewer(palette = "Blues") +   # color scheme for the categories
    ggtitle("Individuals less than 60 years of Age with Breast Cancer Tumor Sizes Larger than 10-14cm ") +
    theme_ipsum()

# Plot
ggplot(lincoln_weather, aes(x = `Mean Temperature [F]`, y = `Month`, fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis(name = "Temp. [F]", option = "C") +
  labs(title = 'Temperatures in Lincoln NE in 2016') +
  theme_ipsum() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    )
```

