---
title: "Bad Plot vs. Good Plot"
author: "Nicole Ayala"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.path= "../Output/")
```

### **Purpose of this Assignment**
- To assess my coding ability ad knowledge of what makes a good plot versus a bad plot.

#### ~ Requirements ~
- All plots must be made in ggplot2, with code visible in the Rmarkdown document.
- The data must be visible.
- The good plot and the bad plot should be based on the same, or similar, data.
- Will be posted on Twitter
- Clearly articulate each principle of data visualization that is broken or not

## BAD PLOT




```{r}
### Load the libraries ###
library(tidyverse)
library(here)
library(ggplot2)
library(magick)
library(devtools)
library(showtext)
library(beyonce)
library(extrafont)
library(dplyr)
library(extrafont)

### Load in the Data ###
cancer<-read.csv(here("GoodPlotBadPlot","Data", "breastcancer_UMC.csv")) # loading data
glimpse(cancer)
# View(cancer)

loadfonts()
# want to output to .ps files instead of .pdf, use:
# loadfonts(device="postscript")
```
```{r}
#font_import(paths = "C:/Users/nikky/AppData/Local/Microsoft/Windows/Fonts")
# Register fonts for Windows bitmap output
font_add(family = "Algerian", regular ="alger.ttf")
showtext_auto()
shrek<-image_read("https://www.pngplay.com/wp-content/uploads/14/Shrek-Family-Transparent-Free-PNG.png")

donkey<-image_read("https://www.pngplay.com/wp-content/uploads/12/Donkey-From-Shrek-Download-Free-PNG.png")
#donkey
#shrek
 
cancer <- cancer %>% 
mutate(class = fct_recode(class, "onions have layers" = "false-recurrence-events")) %>%
  mutate(class = fct_recode(class, "do you know the muffin man" = "recurrence-events")) 
``` 
```{r}
ggplot(data = cancer, 
       aes(x=breast, 
           fill=class)) +
  geom_bar(position='dodge')+
  theme(text = element_text(family = "Algeria")) +
  labs(title='the ploT whith datA?!',
       subtitle ="class versus count, i dunno", 
       x='breast',
       y='count', 
       fill = 'Oh, Shrek. Dont worry.\nThings just seem bad because its dark and rainy\nand Fionas father hired a sleazy hitman to whack you.') +
  
  theme_void() +
  scale_x_discrete(labels = c("donk", "kee")) +
   scale_fill_manual(values = beyonce_palette(8)) +
  theme(plot.title = element_text(size = 95, hjust = 0.5, color ="green3"),
        plot.title.position = "plot", # title adjustments
        plot.subtitle = element_text(size =86,hjust = 0.5, lineheight = 1.1, color = "green2"),
        axis.title = element_text(size=33,color = "grey", hjust = 0.5), # title adjustments
        axis.text = element_text(size = 95,color = "tomato1",hjust = 0.5), # text adjustments
        panel.grid = element_line(linewidth = 0.35, linetype = "dotted", color = "grey"), # background gridlines
        plot.margin = unit(c(1.0, 1.0, 1.0, 1.0), "cm"), # margin adjustments
        plot.background = element_rect(color = "coral1", fill="springgreen1"))

```  
```{r}
# read in the plot as a magick image and make a composite plot
shrekplot<-image_read(here("GoodPlotBadPlot","Output","BadPlot.png"))
shrekout <- image_composite(shrekplot, shrek, offset = "+30+70") 
donkout <- image_composite(shrekout, donkey, offset = "+800+200")
donkout 



```




### Reasons as to why this was a *bad plot*

For starters, there is no clean cut story the data entails. The data was never filtered out so the data itself is not very tidy. But getting to the more aesthetic appeal of this plot, it's not a positive eye catcher but it will turn heads. On a positive note there is a title, subtitle and  axes. However, they are at the worst text size and colors. I apologize in advance to anyone who is red/green colorblind, this is severely difficult to read and interpret. The title and subtitles give no actual meaningful summary, suggestion or overview about what this plot is even about. The background is an unpleasant neon green that is not extremely appealing or pleasing to the eye as well as similar in color ot the title and subtitles. It has a legend that does not correlate to the actual data that was used. The bar plot colors were based off of the beyonce palette, so it was a step up from just keeping it to the original base R colors. However, the size of the plot is tiny meaning it is difficult to read and not clear about what this plot is depicting. It poorly communicates data in a not responsible manner. All in all, I do not think this plot looks awesome in the slightest way. This would definitely not follow reproducible protocol.

## GOOD PLOT

```{r}

### Load the libraries ###

library(tidyverse)
library(here)
library(ggplot2)
library(hrbrthemes)
library(janitor)
library(devtools)
library(dplyr)
library(showtext)
library(extrafont)
library(gridExtra)




### Load in the Data ###
cancer <- read.csv(here("GoodPlotBadPlot","Data", "breastcancer_UMC.csv"))
glimpse(cancer) # look at data
#View(cancer)
attach(cancer) # all the data variables
head(cancer) # shows beginning of  dataset
str(cancer) # shows structure of dataset
#summary(cancer) # shows each column and other attributes


# pair up corr. coefficient btwn variables
#pairs.panels(cancer) # menopause and age is stat sig 
#df <- data.frame(`tumor-size`=c('0-4', '5-9', '10-14', '15-19', '20-24', '25-29', '30-34', '35-39', '40-44', '45-49', '50-54'))


### Filtering the Data ### Part 1
# cancer = filter(cancer, `breast-quad`!= '?' & node.caps != '?')

### Filter Data ### Part 2
data <- cancer %>%
filter(tumor.size>"35-39") 


## Loading Google fonts (https://fonts.google.com/)

#select('tumor_size', age, class, `deg-malig`) %>%
#group_by(class) %>%
#mutate(avg = mean(`tumor-size`, na.rm = TRUE)) %>%
#filter(class == "recurrence-events") %>%
#filter('tumor-size'>"10-14" & age<"60-69")

```
```{r}
# Make the Ridgeline Plot
# Register fonts for Windows bitmap output
font_add(family = "Comic", regular = "comic.ttf")
showtext_auto()
ggplot(data=data,aes(x=`deg.malig`,# degree of malignancy
           group = `tumor.size`,# group by tumor size
           fill = tumor.size)) + #  fill by tumor size
  geom_density(adjust=1.5, alpha=0.5,  show.legend = FALSE) + # adjusting size
 theme(text = element_text(family = "Comic")) +
  theme_ipsum_pub() + # cute rainbow theme
  theme_minimal() +
  labs(x = "Degree of Malignancy", y = "Density", title = "Correlation of the Degree of Malignancy w/ Tumor Size within Patients that Have Breast Cancer", subtitle = "There are 4 ranges of tumor sizes from 40-54 (cm) depicting the larger the mass size of the tumor, the more risk it will be malignant.") + # title and subtitles + x and y axes
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  scale_color_brewer(palette = "Blues") +
   scale_fill_discrete(breaks=c('40-44','45-49', '50-54')) + # arranged order of legend
  theme(axis.title = element_text(size = 30, color = "mediumorchid4"),
  panel.background = element_rect(fill = "white"))
  # background color is white


   

ggsave(here("GoodPlotBadPlot","Output","GoodPlot.png"), # saved the good plot
        width = 7, height = 8) # adjust size of graph in inches

```
### Reasons as to why this was a *good plot*

To begin, the colors follow a theme of shades of pretty pastels within this density plot. The data is more clear as there are not too many variables shown in the legend as well as on the plot. I filtered out the 3 highest ranges of tumor sizes and compared them to its potential degree of malignancy. The data is not difficult to follow as none of the colors are the same and the x axes are properly labeled at a reasonable size. It is also a bit more inclusive when it comes to people with colorblind conditions. The title and subtitle provide meaningful information that will help the individual understand what this plot is showing in regards to breast cancer and risk of a malignant tumor based on its size. I think it's pretty awesome as it's simple and straightforward.